---
title: "Report 1"
author: "Jason Ola"
date: "`r Sys.Date()`"
output: rmdformats::downcute
---

## Import packages

```{r message=FALSE}
library(tidyverse)
library(lubridate)
library(tsibble)
```

## Part 1 : Import dataset

```{r message=FALSE}
cheese_data <- read_csv("data/cheese_data.csv")

cheese_data_f1 <- cheese_data %>% 
  filter(factory == "f1")

cheese_data_f2 <- cheese_data %>% 
  filter(factory == "f2")
```

## Part 2 : Clean the data (manage missing values)

```{r}
sum_na <- cheese_data %>% 
  na_if(0) %>% 
  is.na() %>% 
  sum()
```

1. There are `r sum_na` values to be replaced

```{r}
to_discard <- cheese_data %>% 
  na_if(0) %>% 
  is.na() %>% 
  as_tibble() %>% 
  mutate(sumrow = rowSums(.)) %>% 
  filter(sumrow == 2) %>% 
  nrow()
```

2. There are `r to_discard` time points to be discarded

```{r}
index_replace <- cheese_data %>% 
  na_if(0) %>% 
  is.na() %>% 
  as_tibble() %>% 
  mutate(sumrow = rowSums(.),
         rownum = row_number()) %>% 
  filter(sumrow == 1) %>% 
  pull(rownum)

index_remove <- cheese_data %>% 
  na_if(0) %>% 
  is.na() %>% 
  as_tibble() %>% 
  mutate(sumrow = rowSums(.),
         rownum = row_number()) %>% 
  filter(sumrow == 2) %>% 
  pull(rownum)
```

3. Remove/replace missing values according to manager

```{r}
col_interested <- 4:8
```

```{r message=FALSE, warning=FALSE}
cheese_data_nona <- cheese_data %>% 
  mutate(rownum = row_number()) %>% 
  filter(!rownum %in% index_remove) %>% #remove rows with 2 values
  na_if(0) %>% 
  mutate(meanrow = rowMeans(.[col_interested], na.rm = T) %>% round(1)) %>% 
  rowwise() %>% 
  mutate_at(vars(col_interested),# choose columns of interest
            funs(if_else(is.na(.),meanrow,.)))#replace by meanrow if value is na
```

## Part 3 : Clean the data (manage unexpected values)

```{r message=FALSE}
number_unexpected <- cheese_data_nona %>% 
  filter_at(vars(col_interested),
            any_vars(. >= 65 |. <= 15)) %>% 
  select(col_interested) %>% 
  rowwise() %>% 
  summarise(sum(. >= 65|. <= 15)) %>% 
  pull()
```

1. There are `r number_unexpected` unexpected values in the data.  
Like before I would just replace the unexpected value if there are only 1 by row.  
If there are more by row I would just discard them.

```{r}
cheese_data_nona_nounexp <- cheese_data_nona %>% 
  mutate_at(vars(col_interested),
            #replace unexpected data with mean of m1,m2,m3,m4,m5
            funs(if_else(.>=65,mean(m1,m2,m3,m4,m5),.))) 
```

## Part 4: Compute monitoring statistics for factory f1

1. Storing xbar and range

```{r}
spc_data_f1 <- cheese_data_nona_nounexp %>% 
  filter(factory == "f1") %>% 
  ungroup() %>% 
  mutate(xbar = rowMeans(.[col_interested])) %>% #compute new mean column
  rowwise() %>% 
  mutate(range = max(m1,m2,m3,m4,m5)-min(m1,m2,m3,m4,m5)) %>% 
  select(-c(rownum,meanrow))
```

```{r}
f1_rows <- spc_data_f1 %>% 
  nrow()
```

2. spc_data_f1 has `r f1_rows` rows

```{r}
avg_xbar <- spc_data_f1 %>% 
  ungroup() %>% 
  summarise(mean(xbar)) %>% 
  pull()

avg_range <- spc_data_f1 %>% 
  ungroup() %>% 
  summarise(mean(range)) %>% 
  pull()
```

3. Average of xbar : `r round(avg_xbar,1)`, and average of range : `r round(avg_range,1)`

## Part 5 : Build Shewhart control charts for factory f1

```{r message=FALSE}
constants <- read_csv("data/XbarR_constants.csv")
```

```{r}
d3 <- constants %>% 
  filter(n == 5) %>% 
  pull(D3)

d4 <- constants %>% 
  filter(n == 5) %>% 
  pull(D4)
```


```{r}
lcl <- constants %>% 
  mutate(lcl = d3*avg_range) %>% 
  summarise(lcl = mean(lcl)) %>% 
  pull()

ucl <- constants %>% 
  mutate(ucl = d4*avg_range) %>% 
  summarise(ucl = mean(ucl)) %>% 
  pull()
```

```{r}
r_chart <- spc_data_f1 %>% 
  ggplot(aes(timepoint, range))+
  geom_line()+
  labs(title = "Factory 1 R chart",
       x = "Timepoint",
       y = "Range")+
  theme_minimal()
r_chart
```

```{r}
r_chart <- r_chart+
  geom_hline(yintercept = c(lcl,ucl),
             color = "red")+
  labs(subtitle = "1. With control limits")
r_chart
```
```{r}
r_chart <- r_chart +
  geom_hline(yintercept = avg_range,
             linetype = "dashed")+
  labs(subtitle = "2. With control limits and average range")
r_chart
```
```{r}
lcl_xbar <- constants %>% 
  mutate(lclxbar = avg_xbar - A2 * avg_range) %>% 
  summarise(lclxbar = mean(lclxbar)) %>% 
  pull()

ucl_xbar <- constants %>% 
  mutate(uclxbar = avg_xbar + A2 * avg_range) %>% 
  summarise(uclxbar = mean(uclxbar)) %>% 
  pull()
```


```{r}
x_bar_chart <- spc_data_f1 %>% 
  ggplot(aes(timepoint,xbar))+
  geom_line()+
  geom_hline(yintercept = c(ucl_xbar,lcl_xbar),
             color = "red")+
  geom_hline(yintercept = avg_xbar,
             linetype = "dashed")+
  labs(title = "Factory 1 X bar chart",
       subtitle = "With control limits and average Xbar",
       x = "Timepoint",
       y = "Xbar")+
  theme_minimal()
x_bar_chart
```

## Part 6: Highlight special causes of variations

```{r}
out_of_bounds <- spc_data_f1 %>% 
  filter(xbar > ucl_xbar | xbar < lcl_xbar) # filter out of control limits
```

```{r}
all_lower_ref_value <- function(x) all(x<avg_xbar) # function used with slider
all_above_ref_value <- function(x) all(x>avg_xbar)
```

```{r}
num_lower <- slider::slide_lgl(.x = spc_data_f1$xbar, 
                  .f = all_lower_ref_value, 
                  .after = 9) %>% 
  sum()

num_above <- slider::slide_lgl(.x = spc_data_f1$xbar, 
                  .f = all_above_ref_value, 
                  .after = 9,
                  # complete because there is only 1 and it is last
                  .complete = T) %>% 
  sum()
```

There are `r num_lower` values that have 10 consecutive values below xbar
There are `r num_above` values that have 10 consecutive values above xbar

```{r}
points_with_10_more <- spc_data_f1 %>% 
  ungroup() %>% 
  # check all values to see if they have 9 more values on threshold (avg_bar)
  mutate(consec = slider::slide_lgl(.x = xbar, 
                     # we use lower only because there are no values above
                                    .f = all_lower_ref_value, 
                                    .after = 9)) %>% 
  filter(consec) %>% #filter these values in
  pull(timepoint) # pull the values
```

```{r}
consecutive_times <- (head(points_with_10_more,1)):
                     (tail(points_with_10_more,1)+9)
# get the last timepoint then substract by length
```

```{r}
consecutive_points <- spc_data_f1 %>% 
  filter(timepoint %in% consecutive_times)
```

```{r}
x_bar_chart+
  geom_point(data = out_of_bounds, 
             mapping = aes(x = timepoint, y = xbar),
             color = "red",
             shape = 8)+
  geom_point(data = consecutive_points,
             mapping = aes(x = timepoint, y = xbar),
             color = "red",
             shape = 21,
             size = 4)
```
 
## Part 7 : Build Shewhart control charts for factory f2

```{r}
spc_data_f2 <- cheese_data_nona_nounexp %>% 
  filter(factory == "f2") %>% 
  ungroup() %>% 
  mutate(xbar = rowMeans(.[col_interested])) %>% #compute new mean column
  rowwise() %>% 
  mutate(range = max(m1,m2,m3,m4,m5)-min(m1,m2,m3,m4,m5)) %>% 
  select(-c(rownum,meanrow))
```

```{r}
avg_xbar2 <- spc_data_f2 %>% 
  ungroup() %>% 
  summarise(mean(xbar)) %>% 
  pull()

avg_range2 <- spc_data_f2 %>% 
  ungroup() %>% 
  summarise(mean(range)) %>% 
  pull()
```

```{r}
lcl2 <- constants %>% 
  mutate(lcl = d3*avg_range2) %>% 
  summarise(lcl = mean(lcl)) %>% 
  pull()

ucl2 <- constants %>% 
  mutate(ucl = d4*avg_range2) %>% 
  summarise(ucl = mean(ucl)) %>% 
  pull()
```

```{r}
r_chart2 <- spc_data_f2 %>% 
  ggplot(aes(timepoint, range))+
  geom_line()+
  labs(title = "Factory 2 R chart",
       x = "Timepoint",
       y = "Range")+
  theme_minimal()
r_chart2
```

```{r}
r_chart2 <- r_chart2+
  geom_hline(yintercept = c(lcl2,ucl2),
             color = "red")+
  geom_hline(yintercept = avg_range2,
             linetype = "dashed")+
  labs(subtitle = "With control limits and mean range")
r_chart2
```
```{r}
lcl_xbar2 <- constants %>% 
  mutate(lclxbar = avg_xbar2 - A2 * avg_range2) %>% 
  summarise(lclxbar = mean(lclxbar)) %>% 
  pull()

ucl_xbar2 <- constants %>% 
  mutate(uclxbar = avg_xbar2 + A2 * avg_range2) %>% 
  summarise(uclxbar = mean(uclxbar)) %>% 
  pull()
```

```{r}
x_bar_chart2 <- spc_data_f2 %>% 
  ggplot(aes(timepoint,xbar))+
  geom_line()+
  geom_hline(yintercept = c(ucl_xbar2,lcl_xbar2),
             color = "red")+
  geom_hline(yintercept = avg_xbar2,
             linetype = "dashed")+
  labs(title = "Factory 2 X bar chart",
       subtitle = "With control limits and average Xbar",
       x = "Timepoint",
       y = "Xbar")+
  theme_minimal()
x_bar_chart2
```
```{r}
out_of_bounds2 <- spc_data_f2 %>% 
  filter(xbar > ucl_xbar2 | xbar < lcl_xbar2)
```

```{r}
all_lower_ref_value2 <- function(x) all(x<avg_xbar2) 
all_above_ref_value2 <- function(x) all(x>avg_xbar2)
```

```{r}
num_lower2 <- slider::slide_lgl(.x = spc_data_f2$xbar, 
                  .f = all_lower_ref_value2, 
                  .after = 9) %>% 
  sum()

num_above2 <- slider::slide_lgl(.x = spc_data_f2$xbar, 
                  .f = all_above_ref_value2, 
                  .after = 9,
                  .complete = T) %>% 
  sum()
```

```{r}
points_with_10_more2 <- spc_data_f2 %>% 
  ungroup() %>% 
  # check all values to see if they have 9 more values on threshold (avg_bar)
  mutate(consec = slider::slide_lgl(.x = xbar, 
                     # we use lower only because there are no values above
                                    .f = all_lower_ref_value2, 
                                    .after = 9)) %>% 
  filter(consec) %>% #filter these values in
  pull(timepoint) # pull the values
```

```{r}
first_points <- points_with_10_more2[1:6]
last_points <- points_with_10_more2[7:length(points_with_10_more2)]
```

```{r}
consecutive_times2 <- (head(first_points,1)):
                     (tail(first_points,1)+9)

consecutive_times3 <- (head(last_points,1)):
                     (tail(last_points,1)+9) 
```

```{r}
consecutive_points2 <- spc_data_f2 %>% 
  filter(timepoint %in% consecutive_times2)

consecutive_points3 <- spc_data_f2 %>% 
  filter(timepoint %in% consecutive_times3)
```

```{r}
x_bar_chart2+
  geom_point(data = out_of_bounds2, 
             mapping = aes(x = timepoint, y = xbar),
             color = "red",
             shape = 8)+
  geom_point(data = consecutive_points2,
             mapping = aes(x = timepoint, y = xbar),
             color = "red",
             shape = 21,
             size = 4)+
  geom_point(data = consecutive_points3,
             mapping = aes(x = timepoint, y = xbar),
             color = "red",
             shape = 21,
             size = 4)
```
3. The 2 factory do not have the same pattern at all even though they have almost the same mean. Factory one is more stable whereas factory 2 is spread further away from the mean.

## Part 8 : Redefine monitoring data for factory f2

```{r}
spc_data_f2_week <- spc_data_f2 %>% 
  ungroup() %>% 
  group_by(week) %>% 
  summarise(avg_week = mean(xbar), 
            range_in_week = mean(range)) %>% 
  mutate(moving_range = abs(range_in_week-lag(range_in_week)))
```

```{r}
M <- spc_data_f2_week %>% 
  summarise(mean(avg_week)) %>% 
  pull()

MR_bar <- spc_data_f2_week %>% 
  summarise(mean(moving_range, na.rm = T)) %>% 
  pull()
```

```{r}
lcl_week <- M - 2.66 * MR_bar
ucl_week <- M + 2.66 * MR_bar
```

```{r}
spc_data_f2_week %>% 
  ggplot(aes(week, avg_week))+
  geom_line()+
  geom_hline(yintercept = c(lcl_week,ucl_week),
             color = "red")+
  geom_hline(yintercept = M,
             linetype = "dashed")+
  labs(title = "Control chart for factory 2",
       subtitle = "By week",
       x = "Week",
       y = "Average value")+
  theme_minimal()
```

