---
title: "A story on volcanoes"
author: "Jason Ola"
date: "`r Sys.Date()`"
output: rmdformats::readthedown
---

```{r include=FALSE}
library(tidyverse)
library(sf)
library(magick)
library(DT)
```

## Wow a large volcano !

```{r echo=FALSE}
image_read("data/img/guatemala-central-america-latin-america-volcano-erupting.jpeg") %>% 
  image_scale("1580")
```

```{r include=FALSE}
#load data but it has a github download limit so I save it below in the comments
#volcano_data <- tidytuesdayR::tt_load("2020-05-12")
```

```{r include=FALSE}
# eruptions <- volcano_data$eruptions
# events <- volcano_data$events
# sulfur <- volcano_data$sulfur
# tree_rings <- volcano_data$tree_rings
# volcano <- volcano_data$volcano

```

```{r include=FALSE}
# eruptions %>% write_csv("data/eruptions")
# events %>% write_csv("data/events")
# tree_rings %>% write_csv("data/tree_rings")
# sulfur %>% write_csv("data/sulfur")
# volcano %>% write_csv("data/volcano")

```

```{r include=FALSE}
eruptions <- read_csv("data/eruptions")
events <- read_csv("data/events")
sulfur <- read_csv("data/sulfur")
tree_rings <- read_csv("data/tree_rings")
volcano <- read_csv("data/volcano")%>% 
  mutate(primary_volcano_type = str_replace_all(primary_volcano_type,
                              c("Caldera\\(s\\)" = "Caldera",
                                "Lava cone\\(es\\)" = "Lava cone",
                                "Stratovolcano\\(es\\)" = "Stratovolcano",
                                "Shield\\(s\\)" = "Shield",
                                "Pyroclastic cone\\(s\\)" = "Pyroclastic cone",
                                "Stratovolcano\\?" = "Stratovolcano",
                                "Complex\\(es\\)" = "Complex",
                                "Lava cone\\(s\\)" = "Lava cone",
                                "Lava dome\\(s\\)" = "Lava dome",
                                "Tuff cone\\(s\\)" = "Tuff cone")))
```


```{r include=FALSE}
large_volcanoes <- volcano %>% 
  left_join(eruptions,
            by = c("volcano_number", "latitude", "longitude","volcano_name")) %>% 
  filter(vei >= 5)
```


```{r include=FALSE}
lv_locations <- large_volcanoes %>% 
  st_as_sf(coords = c("longitude","latitude"), crs = 4326)
```

```{r include=FALSE}
world <- st_read("data/world/ne_110m_admin_0_map_units.shp")
```

```{r include=FALSE}
color1 = "#211103"
color2 = "#9F2042"
```

## But... Where are they ?!

### Let's see where the large volcanoes are located !

```{r echo=FALSE}
world %>% filter(SOVEREIGNT != "Antarctica") %>% 
  ggplot()+
  geom_sf(fill = color1,
          color = color1,
          size = 0.5)+
  geom_sf(data = lv_locations,
          aes(size = vei),
          shape = 17,
          color = color2,
          alpha = 0.7)+
  scale_size_continuous(range = c(2,5),breaks = c(5,6,7))+
  labs(title = "Location of large volcanoes",
       subtitle = "with Volcano Explosivity Index \n",
       caption = "Source : github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-05-12")+
  theme_void()+
  theme(legend.position = "bottom",
        legend.title = element_blank(),
        text = element_text(family = "Roboto Slab"),
        plot.title = element_text(hjust = 0.5,
                                  face = "bold",
                                  color = color1),
        plot.subtitle = element_text(hjust = 0.5,
                                     face = "bold",
                                     color = color2),
        plot.caption = element_text(hjust = 0.5,
                                    size = 6))
```

### Here's a table with all volcanoes, their volcano explosivity index and where they are located

###### Volcanic explosivity index (VEI) : Relative measure of the explosiveness of volcanic eruptions

```{r echo=FALSE}
volcano %>% 
  left_join(eruptions,
            by = c("volcano_number", "latitude", "longitude","volcano_name")) %>% 
  select(volcano_name, country,vei) %>% 
  filter(!is.na(vei)) %>% 
  arrange(desc(vei)) %>% 
  distinct(volcano_name, country,vei, .keep_all = F) %>% 
  rename("Volcano" = volcano_name, 
         "Country" = country,
         "VEI" = vei) %>%
  datatable(
    options = list(initComplete = JS("function(settings, json)
                      {$(this.api().table().header()).css(
                      {'font-family' :'Roboto Slab'});}"),
                   dom = "t p"
        )) %>% 
  formatStyle(columns = " ",
    backgroundColor = color2,
    color = color1,
    fontFamily = "Roboto Slab",
    target = "row")
```

### And here are the countries where we can find most of them

```{r echo=FALSE}
volcano %>% 
  count(country) %>% 
  arrange(n) %>% 
  tail(10) %>% 
  mutate(country = fct_inorder(country)) %>% 
  ggplot()+
  geom_col(aes(x = country, 
               y = n),
           fill = color2,
           width = 0.6)+
  labs(title = "Top 10 countries",
       subtitle = "with the most volcanoes",
       caption = "Source : github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-05-12",
       x = "",
       y = "")+
  coord_flip()+
  theme_minimal()+
  theme(text = element_text(family = "Roboto Slab"),
        plot.title = element_text(face = "bold",
                                  color = color1,
                                  hjust = 0.3),
        plot.subtitle = element_text(face = "bold",
                                     color = color2,
                                     hjust = 0.28),
        plot.caption = element_text(size = 6,
                                    hjust = 0.3)
        )
```

### But they are not here by accident...

```{r include=FALSE}
plates <- st_read("data/tectonicplates-master/PB2002_boundaries.shp")
```

```{r include=FALSE}
all_loc <- volcano %>% 
  left_join(eruptions,
            by = c("volcano_number", "latitude", "longitude","volcano_name"))%>% 
  st_as_sf(coords = c("longitude","latitude"), crs = 4326)
```

```{r echo=FALSE}
plates %>% 
  ggplot()+
  geom_sf(color = color1)+
  geom_sf(data = all_loc,
          color = color2,
          shape = 17,
          size = 1)+
  labs(title = "Position of world volcanoes",
       subtitle = "on tectonic plates",
       caption = "Source : github.com/fraxen/tectonicplates 
       github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-05-12")+
  theme_void()+
  theme(text = element_text(family = "Roboto Slab"),
        plot.title = element_text(face = "bold",
                                  color = color1,
                                  hjust = 0.5),
        plot.subtitle = element_text(face = "bold",
                                     color = color2,
                                     hjust = 0.5),
        plot.caption = element_text(hjust = 0.5,
                                    size = 6)
        )
```

We see that the volcanoes are mainly located on the edge of tectonic plates, this is where a lot of seismic activity occurs. Plates go one under the other and create lava resurgences on the surface, thus creating volcanic activity. 

```{r eval=FALSE, include=FALSE}
volcano %>% 
  distinct(primary_volcano_type) %>% 
  arrange(primary_volcano_type)
```
## How explosive are each volcano types in general ?

```{r echo=FALSE, message=FALSE, warning=FALSE}
volcano %>% 
  left_join(eruptions,
            by = c("volcano_number", "latitude", "longitude","volcano_name")) %>% 
  arrange(vei) %>% 
  filter(primary_volcano_type != "Tuff cone") %>% #Tuff cone has only na values
  mutate(primary_volcano_type = fct_inorder(primary_volcano_type)) %>% 
  ggplot()+
  ggridges::geom_density_ridges(aes(x = vei,
                                    y = primary_volcano_type),
                                fill = color2,
                                na.rm = T)+
  labs(title = "Different types of volcanoes",
       subtitle = "and their overall VEI",
       y = "",
       x = "VEI",
       caption = "Source : github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-05-12")+
  theme_minimal()+
  theme(panel.grid.major.y = element_blank(),
        text = element_text(family = "Roboto Slab"),
        plot.title = element_text(hjust = 0.3, 
                                  color = color1,
                                  face = "bold"),
        plot.subtitle = element_text(hjust = 0.3,
                                     color = color2,
                                     face = "bold"),
        axis.title.x = element_text(hjust = 0.4),
        plot.caption = element_text(hjust = 0.3,
                                    size = 6))
```

## What type of volcano scare people most ?

Let's see locally how dangerous a certain volcano is for it's surroundings

```{r echo=FALSE}
large_volcanoes %>% 
  filter(primary_volcano_type != "Subglacial") %>% #not interesting
  group_by(primary_volcano_type) %>% 
  summarise(pop10 = mean(population_within_10_km),
            pop30 = mean(population_within_30_km),
            pop100 = mean(population_within_100_km)) %>% 
  ggplot()+
  geom_point(aes(x = primary_volcano_type,
                 y = 0,
                 color = pop100),
             size = 35)+
  geom_point(aes(x = primary_volcano_type,
                 y = 0,
                 color = pop30),
             size = 20)+
  geom_point(aes(x = primary_volcano_type,
                 y = 0,
                 color = pop10),
             size = 10)+
  geom_point(aes(x = primary_volcano_type,
                 y = 0),
             shape = 17,
             color = color1,
             size = 5)+
  scale_color_gradient(low = color2,
                       high = color1,
                       trans = "log10",
                       breaks = c(1e4,1e5,1e6), 
                       labels = c("High","Med","Low"))+
  labs(title = "Danger zones of large volcanoes (VEI 5-7)",
       subtitle = "according to living population nearby",
       color = "Danger",
       caption = "Source : github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-05-12",
       x = "Zones from center to out : 10, 30 and 100km")+
  guides(color = guide_colorbar(reverse = T))+
  theme_minimal()+
  theme(legend.position = "bottom",
        panel.grid = element_blank(),
        axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.text.x = element_text(vjust = 30,
                                   color = color1),
        text = element_text(family = "Roboto Slab"),
        plot.title = element_text(color = color1,
                                  hjust = 0.5,
                                  face = "bold"),
        plot.subtitle = element_text(color = color2,
                                     hjust = 0.5,
                                     face = "bold"),
        plot.caption = element_text(hjust = 0.5,
                                    size = 6),
        axis.title.x = element_text(vjust = 20,
                                    size = 8,
                                    color = color2))


```





